---
title: "Veteran Population Data for the 41st Congression District"
output: 
  pdf_document:
    toc: true
---

# Veteran Overview

This data is the US census data from 2017 showing current breakdown of Veteran Population in Riverside County.

Students here for UCR for veterans

Major bus route, 91, 215.

Hard to get from temecula to Loma linda. 

```{r echo=FALSE}
library(readxl)
library(kableExtra)
library(ggplot2)
VetOver <- read_excel("Data.xlsx", sheet = "Vet_census", 
    range = "A1:B6")
VetOver <- as.data.frame(VetOver)

VetOver %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

colnames(VetOver) <- c("Catagory", "Number")

ggplot(data=VetOver, aes(x=VetOver$Catagory, y=VetOver$Number, fill=Catagory)) + geom_bar(stat="identity") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) + ylab("Number of Veterans") + xlab("")


```

A trend that it may be increasing. Show a figure of vet population over time. 


```{r echo=FALSE}
library(readxl)
library(kableExtra)
library(ggplot2)

Wait <- read_excel("Data.xlsx", sheet = "VA_waittimes")
Wait <- as.data.frame(Wait[,3:18])
Wait%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

ggplot(data=Wait, aes(x=Wait$Location, y=Wait$`PC Avg Wait Time in Days 13`, fill=Wait$Location)) + geom_bar(stat="identity") + theme(legend.position = "none", axis.text.x = element_text(angle = 90)) + ylab("Primary Care Wait time") + xlab("")

```
History - Jerry Louis 

Capacity -  VA health care locations compared to demographics, kinds of care/specialties
  outpaitent services
  number of patients
  Mental Health
  Community based clinics are based on Health Centers
  Trajectory of capacity
  Ecancer
  telemedicine

Accessiblity - Waittimes, transportation/drive times, San bernadino,

Quality - Complaint data, Ratings, Long beach data from Gerry, Takano may be able to get data for 
  $600 per patient, cost of patient is actualy \$200

Learn more about Arcadia location that just opened- Nepalitono, june, shift, their case for clinic - provides access to those without a car. 

Ideal outcome: Winning contract for 5 CBOCs, over building one here. RFP
Hybrid, community owned, Va owned. CBOCs are lucrative

Email Ignacio about getting VA rank and branch info for va clinics so we can look at the needs of each branch and what will be needed here as far as services go. 


